---
title: "R Notebook"
output: html_notebook
---

# Homework week 11 - day 3 - Clustering

This homework involves the mall_customers.csv data we’ve provided, based on data 
collected in a shopping mall. It will give you an opportunity to practice your new clustering skills.

```{r}
library(tidyverse)

```

# Question 1
Data exploration

```{r}
customers <- read_csv("data/mall_customers.csv")
head(customers)
dim(customers)
skimr::skim(customers)
```

We have 200 observations, with 5 columns: customerID, gender, age, annual income (in k$) and 
spending score (running from 1-100)

So each row represents one customer. Spending score potentially is a score for how
much money is spend in the mall

We have no missing datapoints, distributions look ok for the type of data it contains 
(always positive, so slighlty right skewed)

# Question 2

We are interested in creating a marketing campaign to target customers based on 
their spending score and annual income. Select an appropriate value for k and 
perform k-means clustering to find if there are meaningful clusters in the data to target the customers.

```{r}
library(janitor)
library(fastDummies)
library(broom)
library(tidyverse)
library(factoextra)
```

```{r}
# we are interested in clustering based on spending score and income
# clean names and select only these:

num_customers <- customers %>% 
 # column_to_rownames("Gender") %>% 
  clean_names() %>% 
  select(annual_income_k, spending_score_1_100)

ggplot(num_customers, aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point()

# interesting pattern, roughly  clusters visible!
```

```{r}
# scaling of the variables

customers_scale <- num_customers %>% 
  mutate_if(is.numeric, scale)

customers_scale

customers_scale %>% 
  as_tibble() %>% 
  pivot_longer(cols = c(annual_income_k, spending_score_1_100),
               names_to = "score_income",
               values_to = "value") %>% 
  group_by(score_income) %>% 
  summarise(mean = round(mean(value)),
            sd = sd(value))

# success!
```

check each of the methods of choosing k:

```{r}
# elbow method
fviz_nbclust(customers_scale,
             kmeans,
             method = "wss",
             nstart = 25)
```
I suggest that the elbow starts at 5

```{r}
# silhouette coefficient

fviz_nbclust(customers_scale,
             kmeans,
             method = "silhouette",
             nstart = 25)
```
- 5 seems to be ideal here again..

```{r}
# gap statistic

fviz_nbclust(customers_scale, kmeans, method = "gap_stat")
```
Here one, but after that 5 also still seems reasonable.

Given the results above, I go for 5 clusters
```{r}
clustered_customers <- kmeans(customers_scale,
                              centers = 5,
                              nstart = 25)

clustered_customers
```

```{r}
#some more info on clusters
tidy(clustered_customers, col.names = colnames(customers_scale))
```

```{r}
# augment assigns each cluster number back to the observations

clusters <- augment(clustered_customers, customers) %>% 
  clean_names()

clusters
```

```{r}
# slightly more informative info per cluster:
clusters %>% 
  select(-customer_id) %>% 
  group_by(cluster) %>% 
  summarise(across(.col = is.numeric, .fns = ~ mean(.x)))

```


# Question 3

Visualise the clustering for your chosen value of k.

```{r}
clusters %>% 
  ggplot() +
  aes(x = annual_income_k, y = spending_score_1_100, colour = cluster) +
  geom_point()
```
Based on this graph, we see 5 groups of customers:
Within group of customers that have a high annual income (mean income 86 and 88), there
is a group with a high spending score, and a group with low spending score.

This same pattern is visible for customers with a low annual income (mean 25 and 26k)

Only customers with an average income (55k), sahre a similar spending score with a mean of 49

```{r}
#let's check some demographics:
#gender

clusters %>% 
  ggplot() +
  aes(x = annual_income_k, y = spending_score_1_100, colour = cluster) +
  geom_point() +
  facet_wrap(~ gender)
```
The different genders give the same clustering pattern.

```{r}
clusters %>% 
  ggplot() +
  aes(x = annual_income_k, y = age, colour = cluster) +
  geom_point() 

# grouping stays roughly valid

clusters %>% 
  ggplot() +
  aes(x = spending_score_1_100, y = age, colour = cluster) +
  geom_point() 

# interesting: 
# - for younger customers with high income, no difference in spending score
# - customers with low or middle income, have a spending score independent of age.

```

# Question 4

Do you think the clustering seems a good fit for this data?

Yes, since the grouping of 5 clusters stays +/- present across visualisations of
different variables.

# Question 5

```{r}
clusters %>% 
  ggplot() +
  aes(x = annual_income_k, y = spending_score_1_100, colour = cluster) +
  geom_point() +
  facet_wrap(~ gender) +
  scale_colour_discrete(labels = c("low income, high spending",
                                 "average income and spending",
                                 "low income, low spending",
                                 "high income, high spending",
                                 "high income, low spending"))
```

