---
title: "R Notebook"
output: html_notebook
---

# Question 1
Load the code_clan_tweets.csv data. Find the number of rows, columns, and list all the variable names.

```{r}
library(tidyverse)
tweets <- read_csv("data/code_clan_tweets.csv")

dim(tweets)
names(tweets)
```

#Question 2
Find the total number of favourites (stored in favorite_count) that CodeClan 
tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet 
was quoted (see the is_quote variable).

```{r}
head(tweets)

tweets %>% 
  group_by(is_quote) %>% 
  summarise(total_favorites = sum(favorite_count))

#425
```

#Question 3
Summarise the mean number of retweets (retweet_count), split by the type of 
platform that was used to tweet (source). Again, omit quoted tweets from the mean

```{r}
tweets %>% 
  filter(is_quote == FALSE) %>% 
  group_by(source) %>% 
  summarise(retweets = mean(retweet_count))
```

#Question 4
Count the total number of likes (i.e. total of favorite_count), split by media 
type, and arrange them from most likes to least. Assume that any tweets without 
a listed media type are of type “text”.

```{r}
glimpse(tweets)

tweets %>% 
  mutate(media_type = coalesce(media_type, "text")) %>% 
  group_by(media_type) %>% 
  summarise(total_likes = n()) %>% 
  arrange(desc(total_likes))
  
  
 # group_by(media_type) %>% 
 # summarise(total = sum(favorite_count))
  
# Based on discussion:
  #n() will give the number of rows!!!
  
tweets %>% 
  mutate(media_type = coalesce(media_type, "text")) %>% 
  group_by(media_type) %>% 
  summarise(total_likes = sum(favorite_count)) %>% 
  arrange(desc(total_likes))

```

#Question 5
Find the mean number of characters that a CodeClan tweet contains. You can either 
find a way to count the text in the text variable, or use the variable display_text_width 
which contains a count. Checking both would be a good idea if possible to see if 
they match (but it’s not essential).

```{r}
head(tweets, n=7)

tweets %>% 
  summarise(mean_length = mean(display_text_width))
# = 149.4188


# other method
tweets %>% 
  mutate(length_of_tweets = str_length(text)) %>% 
  summarise(mean_length2 = mean(length_of_tweets))
# = 188.3718

#They do differ!
tweets %>% 
  mutate(length_of_tweets = str_length(text)) %>% 
  select(display_text_width, length_of_tweets)
  
```

#Question 6
The code_clan_info.csv data file contains status_url along with other info. 
Load this in, and join it to the code_clan_tweets tibble, so that you have a 
status_url for each tweet. Decide which variable to join the tibbles on.

```{r}
cc_info <- read_csv("data/code_clan_info.csv")
head(cc_info)

names(tweets)
names(cc_info)

left_join(tweets, cc_info, "tweet_id")
```

#Question 7
From your new joined data, create a new tibble codeclan_hashtags containing only 
tweet_id and hashtags, then convert hashtags into lowercase for analysis. 
Keep only those tweets with hashtags.

```{r}
codeclan_hashtags <- left_join(tweets, cc_info, "tweet_id") %>% 
  select(tweet_id, hashtags)

codeclan_hashtags %>% 
  mutate(hashtags = str_to_lower(hashtags)) %>% 
  drop_na()

# better to do last row as:
filter(!is.na(hashtags))
```
**Extension**

#Question 8
Some tweets have more than one hashtag, denoted by the c( symbols at the start 
of the string. Use the str_sub() and str_detect() functions and your codeclan_hashtags 
tibble from above to find all the cases in which the hashtag string begins with 
charactersc(.

```{r}
codeclan_hashtags %>% 
  mutate(hashtags = str_sub(hashtags, end = 2)) %>% 
  filter(str_detect(hashtags, "c\\("))



```

#Question 9
Use the str_detect() function to find all tweets with text that mentions 
“Edinburgh”, and count how many such tweets there are.

```{r}
tweets %>% 
  select(text) %>% 
  mutate(text = str_to_lower(text)) %>% 
  filter(str_detect(text, "edinburgh")) %>% 
  nrow()
```

#Question 10
Use str_extract_all() with a regular expression to find out which Twitter users 
CodeClan have been tweeting.

```{r}
tweets

tweets %>% 
  select(text) %>%
  str_extract_all("@[a-zA-Z0-9[:punct:]]+")

```


