---
title: "R Notebook"
output: html_notebook
---

# Weekend Homework - Week 06 - Statistics

## Background
You work for an animal conservation organisation and are looking to do some 
investigating into some of the animals to provide evidence for some projects you 
are looking to run.

```{r}
library(tidyverse)
library(ggplot2)
library(skimr)
library(infer)

data(msleep)
```

## 1.1 Hypothesis testing
### Question 1
Explore the dataset and familiarise yourself with it.

```{r}
head(msleep)
skim(msleep)
glimpse(msleep)
```

### Question 2 - 
Jabberwockies sleep for around 7 hours a night, on average. Perform an appropriate 
statistical test to determine whether the mean sleep_total in the sampled population 
of animal types differs from the typical value for jabberwockies

```{r}
# 1. Define our test and significance level
# H0: Mean(sleep_total) = 7 (mean_jabberwockies)
# H1: Mean(sleep_total) ≠ 7

# we use standard α = 0.05

# 2. Calculate test statistic
observed_stat <- msleep %>% 
  summarise(mean_sleep = mean(sleep_total)) %>% 
  pull()
observed_stat #10.43

# 3. we create the null distribution by bootstrapping
null_distribution <- msleep %>% 
  specify(response = sleep_total) %>% 
  hypothesize(null = "point", mu = 7) %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "mean")

# 4. visualise and plot the observed statistic on the null distribution
null_distribution %>% 
  visualise(bins = 25) +
  shade_p_value(obs_stat = observed_stat, direction = "both") #both, because we just want to test for significant difference
```

```{r}
# 5. calculate the p-value
p_value <- null_distribution %>% 
  get_p_value(obs_stat = observed_stat, direction = "both")
p_value
```
p_value = 0.001 (is shown as 0, but cannot be 0).

__Conclusion:__
We reject H0, as P-value < α = 0.001 < 0.05
The average total sleep of all animals is significantly different from the average 
sleep of jabberwockies

### Question 3 - 
Perform an appropriate statistical test to determine whether omnivores sleep for 
significantly longer than herbivores, on average.

```{r}
# 1. create the dataset
vore_sleep <- msleep %>% 
  select(vore, sleep_total) %>% 
  filter(str_detect(vore, "omni|herbi"))

# 2. Define our test and significance level
# H0: Mean(sleep_omnivore) = Mean(sleep_herbivore)
# H1: Mean(sleep_omnivore) > Mean(sleep_herbivore) -> Mean(sleep_omnivore) - Mean(sleep_herbivore) > 0

# we use standard α = 0.05

# 3. create null distribution
null_distribution_vore <- vore_sleep %>% 
  specify(response = sleep_total, explanatory = vore) %>% #the relationship between length and species we test
  hypothesize(null = "independence") %>% # null hypothesis = no relationship
  generate(reps = 5000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("omni", "herbi")) #sample stat is mean(omni)-mean(herbi)

# 4. calculate observed statistic
observed_stat_vore <- vore_sleep %>% 
  specify(response = sleep_total, explanatory = vore) %>% 
  calculate(stat = "diff in means", order = c("omni", "herbi"))

observed_stat_vore

# 5. p-value
p_value_vore <- null_distribution_vore %>% 
  get_p_value(obs_stat = observed_stat_vore, direction = "right")

```
6. conclusion

p-value > α, 0.131 > 0.05
We cannot reject H0, there is no significant difference in total sleep between 
herbivores and omnivores

### Question 4 - 
Perform an appropriate statistical test to determine whether the proportion of 
domesticated animal types in the population of animal types is greater than 5%. 

-> we need to do a one-sample proportion hypothesis test

```{r}
# 1. Define our test and significance level
# H0: Proportion(domesticated_animal_types) = 0.05
# H1: Proportion(domesticated_animal_types) > 0.05

# we use standard α = 0.05
head(msleep)

# 2. edit/create the dataset (lots of NA's for non-domesticated animals)
domestic_msleep <- msleep %>% 
  mutate(is_domestic = if_else(conservation == "domesticated", TRUE, FALSE, 
                               missing = FALSE), .before = "genus")

# 2. create null distribution
null_distribution_prop <- domestic_msleep %>% 
  specify(response = is_domestic, success = "TRUE") %>% 
  hypothesize(null = "point", p = 0.05) %>% 
  generate(reps = 5000, type = "draw") %>% 
  calculate(stat = "prop")

head(null_distribution_prop)

# 3. calculate observed stat
observed_stat_prop <- domestic_msleep %>% 
  group_by(is_domestic) %>% 
  summarise(prop = n()/nrow(domestic_msleep)) %>% 
  filter(is_domestic == "TRUE") %>% 
  pull() # 0.12

# 4. let's visualise
null_distribution_prop %>% 
  visualise(bins = 25) +
  shade_p_value(obs_stat = observed_stat_prop, direction = "right")

# 5. let's calculate the p-value
p_value <- null_distribution_prop %>% 
  get_p_value(obs_stat = observed_stat_prop, direction = "right")

p_value # 0.01
```
Conclusion:
pvalue < α
0.01 < 0.05
We therefore reject H0 
The proportion of domestic animals is significantly higher than 5%

## 1.2 Defining the hypothesis

### Question 1 - 
*-> Test to use:*
One-sample independent hypothesis test for proportion

*-> Defining H0 and H1:*
__H0__ = The proportion of people heard of shop is equal to 40%
  = π(heard_of_shop) = 0.4
  
__H1__ = The proportion of people heard of shop is more than 40%
  = π(heard_of_shop) > 0.4
  
*-> method to generate null distribution:*
  = draw

### Question 2 - 
*-> Test to use:*
two-sample paired hypothesis test for proportions

  (paired because the users, are using the website before and than will either see 
  a edited or un-edited version). Proportions, because you want to compare the 
  proportion of clicks on the banner for each group.)

*-> Defining H0 and H1:*
__H0__ = The proportion of people in group A clicking on the banner is 
          equal to the proportion of people in group B clicking on the banner
  = π(group_A_clicking_on_banner) = π(group_B_clicking_on_banner)
  
__H1__ = The proportion of people in group A clicking on the banner is 
          smaller than the proportion of people in group B clicking on the banner
  = π(group_A_clicking_on_banner) < π(group_B_clicking_on_banner)
  
*-> method to generate null distribution:*
  = permutation
  
### Question 3 - 
*-> Test to use:*
one-sample mean hypothesis test

*-> Defining H0 and H1:*
__H0__ = The width of the produces car part is equal to the standard 145mm
  = µ(part_width) = 145
  
__H1__ = The width of the produces car part is unequal to the standard 145mm
  = µ(part_width) ≠ 145
  
*-> method to generate null distribution:*
  = bootstrap

## 1.3 Interpreting the results

### Question 1
Coffee shop problem. Significance level: 0.05, calculated p-value: 0.07

Conclusion:
p-value > 0.05
We fail to reject H0, we have failed to prove that the proportion of people 
heard of the coffee shop is unequal to 40%.

### Question 2
Website company problem. Significance level: 0.01, p-value: 0.006

Conclusion:
p-value < 0.01
We can therefore reject H0, the proportion of people in group A clicking on the banner is 
significantly smaller than the proportion of people in group B clicking on the banner.

### Question 3
Manufacturing company problem. Significance level: 0.05, p-value: 0.55
          
Conclusion:
p-value > 0.05
We therefore fail to reject H0, the average width of the produced car part is 
statistically significantly equal to the standard 145mm.