---
title: "W1 D2 Homework Thijmen"
output: html_notebook
---

# ##### Question 1
Load in the dataset and have a look at it. What are the dimensions, variable types, 
variable names, etc.?

Loading in the dataset and load packages
```{r}
library(CodeClanData)
library(tidyverse)

drinks_content


dim(drinks_content)

names(drinks_content)
```
check variable types and contents
```{r}
glimpse(drinks_content)
```
check first rows of data
```{r}
head(drinks_content)
```
what is the underlying datatype
```{r}
class(drinks_content)
```

# ####### Question 2
Let’s first investigate the calories of different drinks. Select the variables 
Beverage_category, Beverage, Beverage prep and Calories from your data.
Since we are interested in the calorie content, check if there are any NA values in the data, 
and drop them if there are.

first check if there are missing values in Calories
```{r}
drinks_content %>% 
  mutate(Calories = is.na(Calories))
```
no missing data (all variables are 'FALSE'), just select the columns we want:

```{r}
drinks_content %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) 
```
Note:
you can remove NA values from columns using:
mutate(Calories != "NA")


# ####### Question 3

Create a new variable (column) called calorie_diff, which stores the difference 
between 135 calories (135 calories = 10 cubes of sugar!) and the calories in each drink. 
(hint: you’ll want to subtract 135 from the calories to see which drink have more or less than 10 cups of sugar).

Create a new column which subtracts 135 from calories from each drink
```{r}
drinks_content %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>% 
  mutate(calorie_diff = Calories - 135)
```

# ##### Question 4
Summarise the mean number of calories in each beverage_category. 
Which 3 category of drinks have the most calories? 
Which 3 drinks have the least? Write a small summary of your findings.

Check mean number of calories per beverage category
```{r}
drinks_content %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>% 
  group_by(Beverage_category) %>% 
  summarise(mean_calories = mean(Calories, na.rm = TRUE)) %>% 
  arrange(mean_calories)
```

Check total number of calories per beverage category
```{r}
drinks_content %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>% 
  group_by(Beverage_category) %>% 
  summarise(sum_calories = sum(Calories, na.rm = TRUE)) %>% 
  arrange(sum_calories)
```

In conclusion: the three beverage categories with the highest mean of Calories are:
1) Smoothies, 2) Frappucino Blended coffee, 3) Signature Espresso

 the beverage category with the lowest total number of Calories are:
 1) Coffee, 2) Frappuccino® Light Blended Coffee, 3) Shaken Iced Beverages


# ##### Question 5
Let’s look at this a different way. What is the average number of calories in each Beverage_prep type?

```{r}
drinks_content %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>% 
  group_by(Beverage_prep) %>% 
  summarise(mean_calories_beverage_prep = mean(Calories, na.rm = TRUE)) %>% 
  arrange(mean_calories_beverage_prep)
```

# ##### Question 6
Which Beverage_prep type contains more than the average calories of all drinks?

first calculate the mean of all drinks, than use that to filter the beverage_prep types
```{r}
drinks_content %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>% 
  group_by(Beverage_prep) %>%
  summarise(mean_calories_beverage_prep = mean(Calories, na.rm = TRUE)) %>% 
  filter(mean_calories_beverage_prep >= mean(drinks_content$Calories))

```
So the average amount of calories in all drinks is 193.8719
#also possible to check with
```{r}
drinks_content %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>% 
  summarise(Mean = mean(Calories)) 
```
and there are 5 beverage_prep types with a higher average mean than the total 
mean of all drinks


# ###### Question 7
Which is the best type of coffee drink 
(and by ‘drink’ we mean the combination of Beverage_category, Beverage and Beverage_prep) 
to get if you’re worried about consuming too many calories?

```{r}
drinks_content %>% 
  select(Beverage_category, Beverage, Beverage_prep, Calories) %>% 
  filter(Beverage_category == "Coffee" | Beverage  == "Coffee" | Beverage_prep  == "Coffee") %>% 
  arrange(Calories) %>% 
  head(n=1)
```
The best coffee drink with lowest calorie content is Short, brewed Coffee.
