---
title: "R Notebook"
output: html_notebook
---

# Homework week 5 - day 2 - Thijmen

## Question 1
Load in the libraries you need, and look at the backpack data. 
Take note of variable names and dimensions.
```{r}
library(tidyverse)
library(CodeClanData)

names(backpack)
```
```{r}
dim(backpack)
head(backpack)
view(backpack)
```

## Question 2
What is the distribution of backpack weights that students are carrying? 
Choose an appropriate graph to show the distribution.

```{r}
backpack %>% 
  ggplot(aes(x = backpack_weight)) +
  geom_histogram(binwidth = 1, col = "white") +
  scale_x_continuous(breaks = 1:35)
```
This graph shows the distribution of backpack weights among students. 
Most students have a backpack weight of 10 kg.

# Question 3
Are there differences in backpack weight between sex?

```{r}
backpack %>% 
  ggplot(aes(x = backpack_weight)) +
  geom_histogram(binwidth = 1, col = "white") +
  scale_x_continuous(breaks = 1:35) +
  facet_grid(~ sex) +
  coord_flip()
```
This histogram shows the distribution of backpack weights among students split per gender. 
We can see that in females, most common backpack weight (10 students) is 10 kg, 
while this is 11 kg in male students.
Top bakcpack weight is 35 kg, caried around by a male student.

# Question 4
What is the relationship between body weight and backpack weight?

```{r}
backpack %>% 
  ggplot(aes(x = backpack_weight, y = body_weight)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)

  
```
This scatterplot with linear regression line shows that there is a slight positive relationship
between backpack_weight and body_weight.
Heavier students tend to carry slightly heavier bakcpacks than their lighter students.

* Obviously we should need to statistically test if this is a significant positive relationship!

# Question 5
Is this relationship different between men and women? 
Choose an appropriate plot and interpret your findings.

```{r}
backpack %>% 
  ggplot(aes(x = backpack_weight, y = body_weight)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  facet_grid(~ sex)
```
Breaking up the scatterplot by deviding them by sex shows that this positive relationship 
is slightly stronger among female students. While overall body weight of male students is higher.

* Obviously we should need to statistically test if this is a significant positive relationship!

# Question 6
What proportion of students report back pain?

```{r}
backpack1 <- backpack %>% 
  mutate(back_problems = recode(back_problems,
                               "1" = "Back problem",
                                "0" = "No back problem"))

backpack1 %>% 
ggplot(aes(x = back_problems, fill = back_problems)) +
  geom_bar() +     # use position = "fill" for proportions (stacks can't go higher than 1)
  xlab("Back issues among students") +
  ylab("Percentage") +
  ggtitle("Percentage students with back issues") +
  theme(legend.position = "none") # or within the geom line: legend.position = "none"


# to get the bar stacked on top of each other, leave x empty and use the 'fill = ' argument
#backpack %>% 
#  ggplot(aes(x = "", fill = back_problems)) +
#  geom_bar(position = "fill") +
#  theme_minimal() +
#  labs(y = "proportion",
#       x = "proportion of students with back pain") +
#  scale_y_continuous(labels = scales::percent)
```
Based on the barplot, we can see that the proportion of students with backpain is 32%

```{r}
# or alternatively, and much quicker:
# use after_stat(prop) as Y variable!

backpack1 %>% 
ggplot(aes(x = back_problems, y = after_stat(prop), group = 1)) +
  geom_bar() +    
  xlab("Back issues among students") +
  ylab("Percentage") +
  ggtitle("Percentage students with back issues") +
  scale_y_continuous(labels = scales::percent_format())

```




# Question 7
Is this proportion different between men and women?

```{r}
backpack1 %>% 
ggplot(aes(x = back_problems, y = after_stat(prop), group = 1)) +
  geom_bar(position = "stack") +
  xlab("Back issues among students") +
  ylab("Percentage") +
  ggtitle("Percentage students with back issues") +
  scale_y_continuous(labels = scales::percent_format()) +
  facet_grid(~ sex) 
```
Yes this is different! Among females: ~42% have backpain, while this is ~19% among males!

# Question 8
Now you’ve investigated your data separately, create plots to describe the 
relationship between backpack weight, body weight, gender and reported backpain.

Try both backpack_weight and body_weight as well as the ratio.

Hint: faceting would be useful here.

```{r}
backpack1 %>% 
  ggplot(aes(x = backpack_weight, y = body_weight, group = back_problems)) +
  geom_point() +
  facet_grid(back_problems ~ sex) +
  geom_smooth(method = lm, se = FALSE) 

# scatterplot shows that there is a slight positive relation between body_weight 
# and back_weight among females with and those without back_problems. 
# This relation is slightly stronger among females without back problems.

# However, 
# scatterplot shows that there is a slight negative relation between body_weight 
# and back_weight among males with back_problems. While a slight positive relation exists 
# among males without back problems.

# Obviously we should statistically test if this is a significant relationship!

```
## Extension
## Question 1
Does reported backpain differ across the year groups? Ignore year 0 in your data.

```{r}
backpack1 %>% 
  filter(year != 0) %>% 
  ggplot(aes(x = back_problems)) +
  geom_bar() +
  facet_wrap(~ year) +
  ggtitle("reported back pain per school year") +
  scale_y_continuous(breaks = 1:20)
  

# Yes, there is a large variation in reported back pain per year. 
# Varying from no reported back pain in year 6 to almost 50% of students with back pain in year 3 with 
# 12 reported cases vs 15 no_back pain cases.
```

## Question 2
Is there a difference between how heavy the bags those in different years carry?
```{r}
backpack %>% 
   filter(year != 0) %>%
  ggplot(aes(x = backpack_weight)) +
  geom_histogram(binwidth = 1, col = "white") +
  facet_wrap(~ year) +
  ggtitle("backpack weight distribution divided per year") +
  scale_x_continuous(breaks = c(1, 5, 10, 15, 20, 25, 30, 35))

```

