---
title: "R Notebook"
output: html_notebook
---

# Homework review 

# Question 5
Write a function that pre-appends "The answer is: " to any character input, while 
for numeric input, it either just returns the number for numbers less than 10, or 
divides the number by 10 for numbers greater than 10. For anything else return "Invalid input".

What is going on?

input -> function -> output
"fourteen" ---> "The answer is fourteen"
"14" ---> "The answer is 14"

20 ---> 2 (if input is >10 --> input/10)
5 ----> 5 (if input <10 --> input)

10 --> "invalid input"
TRUE --> "Invalid input"

```{r}
library(tidyverse)
```


```{r}
input_value <- 14

# case 1: character input --> "The answer is _"

if(is.character(input_value)) {
 str_c("The answer is", input_value, sep = " ")
}

if(input_value > 10) {
  input_value/10
}

if(input_value < 10){
  input_value
}



#
#10 --> "invalid input"
#TRUE --> "Invalid input"

```

function starts

test if whether character or not

test if character is number
  test if divisible by 10

   -> yes --> /

   -> no --> x

else
  -> invalid

```{r}
print_answer <- function(input_value) {
  
  if(is.character(input_value)) {
    return_value <- str_c("The answer is", input_value, sep = " ")
  } else if (is.numeric(input_value)) {
    if(input_value > 10) {
      return_value <- input_value /10
    }else if (input_value <10) {
      return_value <- input_value
    }else{
      return_value <- "Invalid input"
    }
    
  } else {
    return_value <- "Invalid input"
  }
  return_value
}

print_answer(14)
print_answer("fourteen")
print_answer(10)
```

