---
title: "R Notebook"
output: html_notebook
---

# Recap day 5 

## Functions

```{r}
do_something <- function(what_does_our_function_take_in) {
  # what the function does
  
  # what the function spits out
  return()
}
```

Functions are mini machines. They take in some stuff, do something to it and give out
something

For example: a function that adds: "David says..." to any input

```{r}
david_says <- function(any_input) {
  return(paste("David says...", any_input))
}

david_says(any_input = "functions are good")
david_says(14)
```

What would we actually use these for?
- long/ often repeated tasks

e.g.
adds one [011000000100101010101111000000110]
multiplier ( ..... adds one x times)
...
...
...

an example, we often copy/paste the following code

```{r}
library(tidyverse)

iris %>% 
  summarise(across(.fns = ~sum(is.na(.x))))

mtcars %>% 
  summarise(across(.fns = ~sum(is.na(.x))))

#instead of copying, with a function we can automate this

count_na <- function(data) {
  data %>% 
    summarise(across(.fns = ~sum(is.na(.x))))
}

iris %>% 
  count_na()
```

We want to add more functionality to our function. We want to give users the chance 
to say whether they only want numeric columns to be summarised across

sometimes

```{r}
#all columns
iris %>% 
  summarise(across(everything(), .fns = ~sum(is.na(.x))))

# just the numeric columns
iris %>% 
  summarise(across(where(is.numeric), .fns = ~sum(is.na(.x))))
```

```{r}
# adding a default argument with param = argument
count_na <- function(data, only_numeric_columns = FALSE) {
  
  
  if (only_numeric_columns) {
    # if they only want nas of numerics
    na_s <- data %>% 
      summarise(across(where(is.numeric), .fns = ~sum(is.na(.x))))
  } else {
    # otherwise return every column
    na_s <- data %>% 
      summarise(across(.fns = ~sum(is.na(.x))))
  }

  return(na_s)
}


iris %>% 
  count_na(only_numeric_columns = TRUE)
```

Refactoring / Rewriting a nested function
Question 5
Write a function that pre-appends “The answer is:” to any character input
while for numeric input, it either just returns the number for numbers less than 10, or divides the number by 10 for numbers greater than 10.
For anything else return “Invalid input”.

```{r}
library(stringr)
```


How could this function be improved?
Some extra things that might help

-> early returns
-> large functions, go through step by step, what the logic is.
ideally write pseudocode.

* before writing a function, write a step-by-step guide
-> if the input is character
    (do that things)
-> if tinoput is number
    (do this)
-> otherwise
    (INVALID)


```{r}
print_answer <- function(input_value) {
 
  #guard clause (guarding against non character, non numeric inputs)
 if ((!is.character(input_value)) & (!is.numeric(input_value))) {
   return("invalid input")
 }
  # only characters and numerics are reaching this point
  
   # if the input is a character....
  if (is.character(input_value)) {
    return_value <- str_c("The answer is", input_value, sep = " ")
  } else if (is.numeric(input_value)) {
    # else if the input is a number...
    if (input_value > 10) {
     #if greater than 10
       return_value <- input_value/10
    } else if (input_value < 10) {
      # if less than 10
      return_value <- input_value
    } else {
      #otherwise
      return_value <- "Invalid input"
    }
  } else {
    return_value <- "Invalid input"
  }
  
  return_value
  
}

print_answer(10)
```

CUrrent less-good things about our function. 
1. We assign invalid input in two cases (value is 10, or value isn't a character or number)
2. Nesting (if's inside of an if)

```{r}
print_answer <- function(input_value) {
 
  # if the input is a character....
  if (is.character(input_value)) {
    return_value <- str_c("The answer is", input_value, sep = " ")
  } else if (is.numeric(input_value)) {
    # else if the input is a number
    if (input_value > 10) {
      return_value <- input_value/10
    } else if (input_value < 10) {
      return_value <- input_value
    } else {
      return_value <- "Invalid input"
    }
  } else {
    return_value <- "Invalid input"
  }
  
  return_value
  
}

print_answer(10)
```

