---
title: "R Notebook"
output: html_notebook
---

# Weekend homework wk01 _ Thijmen

Data: Goodreads - book reviews and reading data

# Start by exploring the data
```{r}
library(tidyverse)

books <- read_csv("data/books.csv")

dim(books) #11123 rows, 13 columns
names(books)
glimpse(books)
view(books)
head(books)
```

Questions of interest to answer:
1 - What are the top rated and worst rated books, authors, publishers etc.?
2 - Are books rated in GoodReads mostly english?
3 - Do more recently published books receive a higher score than older books?
4 - Do books with a text-review receive a higher average rating than books without?
5 - We don't like transphobes, so lets change all rates of JK Rowling to 0

# Lets first get a better picture of missing data
```{r}
books %>% 
  summarise(across(.fns = ~ 
                     sum(is.na(.x))
                   ))

#no missing data!
```
# Question 1
* What are the top rated and worst rated books, authors, publishers etc.?

```{r}
# Authors
books %>% 
  select(title, authors, average_rating, ratings_count, text_reviews_count) %>% 
  group_by(authors) %>% 
  summarise(mean_rating = mean(average_rating)) %>% 
  slice_max(mean_rating, n = 10)
```
In total 18 authors have an average rating of 5 (maximum score)

Let's check how many people have actually rated these authors:
```{r}
books %>% 
  select(title, authors, average_rating, ratings_count, text_reviews_count) %>% 
  group_by(authors) %>% 
  summarise(mean_rating = mean(average_rating), 
            total_counts = sum(ratings_count)) %>% 
             slice_max(mean_rating, n = 10)
```
So most authors that have highest 'mean_rating' have 0, 1 or 2 counts. 
Let's filter out all books with <25 ratings_count and put that in a variable 
'books_min25_rating'
```{r}
books_min25_rating <- books %>% 
  select(title, authors, average_rating, ratings_count, text_reviews_count) %>% 
  filter(ratings_count >= 25)

books_min25_rating %>% 
  arrange(ratings_count) # success!
```
Let's check top rated authors again (highest average scoring rating)
```{r}
books_min25_rating %>% 
  group_by(authors) %>% 
  summarise(mean_rating = mean(average_rating)) %>% 
  slice_max(mean_rating, n = 10)

# Top rated author(s) Saul Leiter/Martin Harrison
  
```

* Do the same for books (we do not need to calculate the average just order on average rating)

```{r}
books_min25_rating %>% 
  arrange(desc(average_rating)) %>% 
  head(n = 5) %>% 
  select(title, authors, average_rating)

# Top rated book: The Complete Calvin and Hobbes, Bill Watterson, 4.82
```
* Do the same for publishers

```{r}
books %>% 
  select(title, authors, average_rating, ratings_count, text_reviews_count, publisher) %>% 
  filter(ratings_count >= 25) %>% 
  slice_max(average_rating, n = 5) %>% 
  select(publisher) %>% 
  head(n = 5)

# Top rated publisher is Andrews McMeel Publishing
  
```

# Question 2
* Are books rated in GoodReads, mostly english?

```{r}
# first lets count and drop all books with ratings_count 0
books %>% 
 select(ratings_count) %>% 
  filter(ratings_count == 0)  %>% 
  nrow() # 80! Lets drop these

books_min1_rating <- books %>% 
  select(title, authors, language_code, average_rating, ratings_count, publication_date, 
         text_reviews_count, publisher) %>% 
  filter(ratings_count > 0)

#lets check how many language groups exist, and how many books are listed per language
books_min1_rating %>% 
  select(language_code, average_rating, ratings_count, text_reviews_count) %>% 
  mutate(count = 1,
         .after = language_code) %>% 
  group_by(language_code) %>% 
  summarise(total_books_per_language = sum(count))

# In total 26 language codes exist, with a minimal rating count of 1

# 4 different codes exist for english, lets minimise that to a single code and save the 
# overview to a variable
totalbooks_language <- books_min1_rating %>% 
  select(language_code, average_rating, ratings_count, text_reviews_count) %>%
  mutate(
    language_code = recode(
      language_code, 
      "en-CA" = "eng", "en-GB" = "eng", "en-US" = "eng" 
    )
    ) %>% 
  mutate(count = 1,
         .after = language_code) %>% 
  group_by(language_code) %>% 
  summarise(total_books_per_language = sum(count))

totalbooks_language

# Let's now count the percentage of books per language of the total
totalbooks <- totalbooks_language %>% 
  summarise(sum(total_books_per_language)) %>% 
  pull()

totalbooks_language %>% 
  mutate(percentage_of_total = total_books_per_language/totalbooks*100,
         .before = total_books_per_language) %>% 
  slice_max(percentage_of_total, n = 5)

# English make up 94.8% of the total nr of books! Biased list..


```

# Question 3
Do more recently published books receive a higher score than older books?
```{r}
# We start again with the min1_rating dataset
books_min1_rating %>% 
  select(title, authors, publication_date, average_rating) %>% 
  # lets focus on years only
  mutate(year = format(as.Date(books_min1_rating$publication_date,
                               format="%m/%d/%Y"), "%Y"),
                       .before = publication_date) %>% 
# Let's just create 2 groups, before 2000 is 'old' and younger than 2000 'new'
mutate(age = case_when(
  year <2000 ~ "old",
  year >=2000 ~ "young"
), .after = year) %>% 
 # filter(is.na(year)) # there are 2 books that have NA for year, but publication date is like normal?!
  group_by(age) %>% 
  summarise(average_rating = mean(average_rating))
  
# In conclusion: 'old' books have an average rating of 3.95, 'young' 3.93
# old win!
```

#Question 4
4 - Do books with a text-review receive a higher average rating than books without?
```{r}
# How many text_reviews are written, and how many ratings are given?
books_min1_rating %>%  #tibble only including book ratings with minimal rate count of 1
  select(title, authors, average_rating, ratings_count, text_reviews_count) %>% 
  summarise(across(where(is.numeric), sum))
# 6029199 text reviews in total, 199578299 ratings in total

books_min1_rating %>%  
  select(title, authors, average_rating, ratings_count, text_reviews_count) %>%
  mutate(text_review_presence = if_else(text_reviews_count > 1, TRUE, FALSE)) %>% 
  group_by(text_review_presence) %>% 
 # filter(text_review_presence == TRUE) %>%  # We don't need this if using group_by
  summarise(average_rate_with_text = mean(average_rating))

#Books without a text review get on average a lower score than those with a text review
# people that like a book are potentially more willing to invest time in writing a review
# but tiny difference!

```

#Question 5
5 - We don't like transphobes, so lets change all rates of JK Rowling to 0
```{r}
books %>% 
  mutate(average_rating = if_else(stringr::str_detect(authors, "J.K. Rowling"),
                                  0, average_rating)
         ) %>% 
  group_by(authors) %>% 
  filter(stringr::str_detect(authors, "J.K. Rowling")) %>% 
  summarise(average_rating = mean(average_rating))
#All books (co-) authored by J.K. Rowling have a average_rating of 0: Success!
```

