---
title: "R Notebook"
output: html_notebook
---

# Question 1 - Load the data, clean_names() and explore it.

```{r}
library(tidyverse)
library(skimr)

phones <- read_csv("data/20190928-items.csv") %>% 
  janitor::clean_names()

head(phones)
skim(phones)

phones %>% 
  ggplot() +
  aes(x = rating) +
  geom_histogram(col = "white")
```
We have data on mobile phone models, including a rating score, total number of reviews given
and price.

Based on the skim() function, total reviews seems to be strongly right skewed. 
Based on skim() and the plot, the review data follows a more normal distribution but 
seems moderately skewed to the left. 

#  Question 2 - Find the brand with the highest numbers of phones in the dataset
```{r}
phones %>% 
  select(brand) %>% 
  group_by(brand) %>% 
  summarise(total_phones = n()) %>% 
  arrange(desc(total_phones))

# samsung has the most phones in the dataset with 397
```

# Question 3 - 
For your top brand, plot the distribution of phone ratings as a probability density, 
overlaying a fitted normal distribution. Do you think the normal distribution provides 
a good model of these ratings?

```{r}
# first just a histogram
phones %>% 
  filter(brand == "Samsung") %>% 
  ggplot() +
  aes(x = rating) +
  geom_histogram()

# lets plot the distribution of ratings as a probability density, overlaying a fitted normal distribution

#calculate sd, and mean:
phone_stats <- phones %>% 
  filter(brand == "Samsung") %>% 
  summarise(
    num = n(),
    mean = mean(rating),
    sd = sd(rating)
  )

phones %>% 
  filter(brand == "Samsung") %>% 
  ggplot(aes(x = rating)) +
  geom_histogram(aes(y = after_stat(density)),
                 col = "white") +
  stat_function(
    fun = dnorm,
    args = list(
      mean = phone_stats$mean,
      sd = phone_stats$sd
    ),
    col = "red"
  )
```
I believe that the normal distribution does provide an ok model for the ratings. Because:
- We see that the data is unimodal spread
- Its highest peak, does overlap with the peak of the normal distribution curve, eventhough 
  two bins particularly have a higher density than the rest. But this equals out with near bins being lower.
- Data shows some left skewed spread, due to an outlier, but this is somewhat compensated by 
  a higher peak on the left hand side of the curve.
  
# Question 4 
We can be more quantitative in comparing the distribution of top brand ratings 
with a normal distribution. You calculated the mean() and sd() of your distribution in 
an earlier question. Use these values to calculate the proportions of ratings 
within one-, two- and three standard deviations of the mean. Compare these proportions 
with those you expect for a normal distribution.

```{r}
# Mean of Samsung rating:
phone_stats$mean # 3.5733
# sd of Samsung rating:
phone_stats$sd #0.6847859

# total samsung ratings:
phones %>% 
  filter(brand == "Samsung") %>% 
  nrow() # 397

# get all ratings 1SD of the mean:
  phones %>% 
  filter(brand == "Samsung") %>% 
  filter(rating >= 3.5733 - 0.6847859) %>% 
  filter(rating <= 3.5733 + 0.6847859) %>% 
  summarise(proportion_in_1sd = n() / 397)
  # 0.7934509	
  
# get all ratings 2SD of the mean:
  phones %>% 
  filter(brand == "Samsung") %>% 
  filter(rating >= 3.5733 - (2*0.6847859)) %>% 
  filter(rating <= 3.5733 + (2*0.6847859)) %>% 
  summarise(proportion_in_1sd = n() / 397)
  # 0.9193955	
  
# get all ratings 3sd of the mean:
  phones %>% 
  filter(brand == "Samsung") %>% 
  filter(rating >= 3.5733 - (3*0.6847859)) %>% 
  filter(rating <= 3.5733 + (3*0.6847859)) %>% 
  summarise(proportion_in_1sd = n() / 397)
  #0.9798489	
  
# Expected proportions of a normal distribution are:
  #1sd:
  pnorm(1) - pnorm(-1) #0.6826895
  # 2 sd:
  pnorm(2) - pnorm(-2) #0.9544997
  # 3 sd:
  pnorm(3) - pnorm(-3) #0.9973002
  
# The samsung ratings do not strictly follow a normal distribution if we look at the 
  # proportions of ratings within 1-2 and 3 SDs of the mean. 
  # The proportion of values of our data are slightly off, with 80% of the data within the first SD of the mean
  # while this is 68% in actual normally distributed data. (similar for 2SD and 3 SD). 
  
  # In conclusion:
  # A large proportion of our data lay outside the actual range of data with a normal distribution!
```

# Extension
```{r}
 phones %>% 
  filter(brand == "Samsung") %>% 
  qqplot()
```

