---
title: "R Notebook - Homework Week01_Day3_ThijmenB"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```



# Homework

# Question 1
After you’ve read in the data and assigned it to a variable with a useful name, 
investigate it. Write down some ideas for what you could do to further investigate/clean 
this data using what you’ve learned today (no need to code it though, just jot down 
thoughts and ideas). Feel free to also write down any general thoughts you get looking 
at the data - this task is just to get you thinking about data and data cleaning/prepping.

```{r, message = FALSE}
# open packages
library(tidyverse)
```

```{r}
# read in the data
movies <- read_csv("data/movies.csv") 

# First get a sense of the data
dim(movies)

names(movies)

head(movies)
```
Based on the first view of the data there are a few things we could do to make the 
dataset more manageable.

1. Column 'overview' seems just filled with a long description: remove for analyses
2. Column 'budget' gives costs, I would change the values into something sensible 
  and readable. The same applies for column 'revenue'
3. Column 'original_title' and 'title' seem to contain the same data: check if true 
  and delete one if this is the case.
4. check if column 'original_language' is identical for all rows ('en')
  if so: delete.
5. Further columns like 'homepage' and 'tagline' do not seem to contain data you 
  can analyse. So perhaps delete if not needed.
  
# Question 2
A lot of interesting stuff in there! But for now, we are only interested in the 
movies’ titles, their runtime as well as their budget. Select the three columns 
that contain this information and save them into a new variable called movies_selected. 
Hint: you will have to make a choice between original_title and title.


```{r}
movies_selected <- movies %>% 
  select(title, runtime, budget)
```

# Question 3
Ok, now take your subsetted data movies_selected, and count the number of missing values in each column.

```{r}
movies_selected %>% 
  summarise(total_missing_title = sum(is.na(title)),
            total_missing_runtime = sum(is.na(runtime)),
            total_missing_budget = sum(is.na(budget)))
```
So only runtime is missing 2 datapoints.

# Question 4
There’s not too many NAs, but there are actually quite a few movies with a 
runtime of 0 in this dataset. Since it’s impossible for a movie to have a 
runtime of 0, use the function na_if to convert runtimes of length 0 into NAs. 
Save this into a new object called movies_runtime. Then count how many missing 
values there are now in the column runtime.

```{r}
movies_runtime <- movies_selected %>% 
  mutate(runtime = na_if(runtime, 0)
  )
movies_runtime
# changed 0s into NA, now count number of NAs
movies_runtime %>% 
  summarise(total_missing_runtime = sum(is.na(runtime))
            )

# in total 37 are missing a value

```

# Question 5
Impute the missing values in runtime with the median runtime using coalesce and 
save this into a new object called movies_imputed. There should now be no missing 
values in the column runtime - check this to make sure.


```{r}
movies_imputed <- movies_runtime %>% 
  mutate(runtime = coalesce(runtime, median(runtime, na.rm = TRUE)))

movies_imputed

movies_imputed %>% 
  summarise(movies_runtime_missing = sum(is.na(runtime)))

# Indeed, no NAs remaining!
```

# Question 6
Finally, use the slice_min and slice_max functions on movies_imputed to get the 
movies with the 10 shortest and 10 longest runtimes.

```{r}
# 10 movies with longest runtimes
movies_imputed %>% 
  slice_max(runtime, n = 10)

# 10 movies with shortest runtimes
movies_imputed %>% 
  slice_min(runtime, n = 10)

```

# Question 7
Using the if_else function, impute anything with a budget below $100 with the 
median budget. Overwrite your movies_imputed so it contains the new budget values

```{r}

movies_imputed_1 <- movies_imputed %>% 
  mutate(budget = if_else(budget < 100, 
                                    median(budget), 
                                    budget))

movies_imputed <- movies_imputed_1 

# check if done correctly
movies_imputed %>% 
  slice_min(budget, n = 10) 

# Yeah!



```


# Extension Q1
Using case_when, create a new column called budget_type and label budgets as follows:

Budgets below 12 million as ‘Small budget’
Budgets between 12 and 40 million as ‘Medium budget’
Budgets 40 millions and above as ‘Big budget’
Save this into a new object called movie_budgets.

```{r}
movie_budgets <- movies_imputed %>% 
  mutate(budget_type = case_when(
    budget < 12000000 ~ "Small budget",
    budget > 12000000 & budget < 40000000 ~ "Medium budget",
    budget > 40000000 ~ "Big budget")
  )

# the "&" part of the second line is not needed, because its going through it 
# based on presence "small budget" exists so it will not run anything over.
# so swecond line should be: "budget < 40000000 ~ "Medium budget"

movie_budgets

#Yeah
```

# Extension Q2
Take the original dataset with all the variables. Using across and where, 
summarise the number of missing values, first across all columns of type 
character, and then across all columns of type numeric.


```{r}
movies %>% 
  summarise(is.na(across(where(is.numeric), sum)))

#?? How??

# Answers during class review:

movies %>% 
  summarise(across(
    .cols = everything(), #here search across all columns
    .fns = ~sum(is.na(.x))
  ))

movies %>% 
  summarise(across(
    .cols = where(is.character), #here search across all character columns
    .fns = ~sum(is.na(.x))
  ))

# and for numeric
movies %>% 
  summarise(across(
    .cols = where(is.numeric), #here search across all character columns
    .fns = ~sum(is.na(.x))
  ))
```


