---
title: "R Notebook"
output: html_notebook
---

# Homework Week 6_day 03

##Q1
Load the data again, clean_names(), and re-familiarise yourself with it
```{r}
library(tidyverse)
library(infer)
library(janitor)

ames <- read_csv("data/ames.csv") %>% 
  clean_names()
```
```{r}
head(ames)
glimpse(ames)
```
##Q2
Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?
```{r}
ames %>% 
  ggplot() +
  aes(x = lot_area) +
  geom_histogram(colour = "white")

#The distribution of lot_area is very narrow, with only 6 bins filled with data
# large majority of the datapoints within a single bin. 
# Very hard to judge if we can call this data normally distributed!

ames %>%
  ggplot(aes(x = lot_area)) +
  geom_boxplot()

#based on the boxplot you see that the data is strongly right skewed.
```

##Q3
Compute and visualise a bootstrap sampling distribution for the mean(lot_area) of the sold houses

```{r}
bootstrap_sampling <- ames %>% 
  specify(response = lot_area) %>%  # specify the variable of interest
  generate(reps = 5000, type = "bootstrap") %>% # generate the replicate resamples
  calculate(stat = "mean") # calculate the statistic of interest for each resample
  
bootstrap_sampling

# visualise
bootstrap_sampling %>% 
  visualise(bins = 30) 

```

##Q4
Use your bootstrap distribution to calculate a 95%
 CI for mean(lot_area), and visualise it on the distribution
 
```{r}
infer_ci_95 <- bootstrap_sampling %>% 
  get_ci(level = 0.95, type = "percentile")

# add to visualisation
bootstrap_sampling %>% 
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = infer_ci_95)

infer_ci_95
```
 
## Q5
You would like to know the mean(lot_area) of the sold houses with higher confidence. Calculate the 99%
 CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the 95%
CI? Does that make sense?

```{r}
infer_ci_99 <- bootstrap_sampling %>% 
  get_ci(level = 0.99, type = "percentile")

infer_ci_95
infer_ci_99
```
infer_ci_95 -> [9877, 10450]
infer_ci_99 -> [9814, 10535]
The 99% CI is broader than the 95% CI, this does make sense!

With 99%, you increase the area you want your sample points to fall in. 
This is wider with 99% of your data, compared to 95% of your data (this is a narrower window)

"It is broader than the 95%
 CI. We haven’t increased sample size, so the only way to increase confidence is 
 by increasing the width of the interval. Again, by analogy: to be more confident 
 of catching the fish, you need to use a bigger net."

##Q6
Calculate the _point estimate_ of the mean(lot_area)

```{r}
p_estimate_mean_lot_area <- bootstrap_sampling %>% 
  summarise(mean = mean(stat))

p_estimate_mean_lot_area
```
-> point estimate of the mean(lot_area) -> 10149.66


##Q1 extension
```{r}

```

# What is CI
-> Statistic is a calculated stat
-> parameter is a 