---
title: "R Notebook"
output: html_notebook
---

# Homework Week1 - Day4

Extra `dplyr` practice

## Question 1
Read in the data and investigate it, e.g. variables, variable types, dimensions etc.


```{r}
library(tidyverse)

wine <- read_csv("data/wine_ratings.csv")

names(wine)
dim(wine)
glimpse(wine)
head(wine)

```

## Question 2
Select columns title, taster_name, points, price, and country, then drop all NAs. 
Save into a variable called wine_subset. Use this variable for all subsequent tasks.

```{r}
wine_subset <- wine %>%
  select(title, taster_name, points, price, country) %>% 
  drop_na()

# Check if correct:
wine_subset %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
#0s across the board!
```

## Question 3
Find the 5 countries with the highest mean points

```{r}
# wine_subset %>% 

wine_subset %>% 
  group_by(country) %>% 
  summarise(mean_country = mean(points)) %>% 
  slice_max(mean_country, n=5)
  
# best scoring countries are England, Austria, Germany, Hungary, Switzerland
  
```

## Question 4
Find the 5 countries with the lowest mean prices

```{r}
wine_subset %>% 
  group_by(country) %>% 
  summarise(mean_country = mean(points)) %>% 
  slice_min(mean_country, n=5)

#worst scoring using the mean is: Brazil, Peru, Mexico, Chile, Lebanon
```

## Question 5
Does the result differ if you use the median instead of the mean for Questions 3 and 4?

- lets find out!

```{r}
wine_subset %>% 
  group_by(country) %>% 
  summarise(median_country = median(points)) %>% 
  slice_max(median_country, n=5)

# Yes, for median on 5th place is Australia instead of Switzerland for mean

wine_subset %>% 
  group_by(country) %>% 
  summarise(median_country = median(points)) %>% 
  slice_min(median_country, n=5)

# There are 5 lowest with the same lowest score: Brazil, Peru, Argentina (different!), 
# Greece (different), and on shared 5th place: Chile, Lebanon, Romania, Morocco, Moldova, Mexico
# So big difference with using the mean here!

```

## Question 6
Find the names of all the tasters using distinct. Then take a random sample of 5 
names using sample_n. Save these 5 names into a new variable called taster_sample 
which should be a character vector containing only the names.

```{r}
taster_sample <- wine_subset %>% 
  distinct(taster_name) %>% 
  sample_n(size = 5) %>% 
  pull()

#pull() lets you save a character list as vector. Vector is a one-dimensiional'list' of things.

#yes, character vector of a list 'type'
  
```
## Question 7
Do the same thing for country, i.e. create a character vector called country_sample 
using distinct and pull, but this time use sample_frac to pick 10 countries at random.

```{r}
#  sample_frac(n = 10) sample_frac takes a fraction of rows to select.
# So lets first check how many countries there are listed
wine_subset %>% 
   distinct(country) %>% 
  nrow()
#So 32 countries listed
#We only want 10 for sample_frac so we calculate:
10/32 #3.2

country_sample <-  wine_subset %>% 
  distinct(country) %>% 
  sample_frac(size = 0.3125, replace = FALSE) %>% 
  pull()

country_sample

```
## Question 8
Filter the rows of wine_subset so you are left only with data from the 5 tasters 
in your taster_sample and from the countries in your country_sample. Save this 
into a new variable called wine_sample.

```{r}

wine_sample <- wine_subset %>% 
  filter(taster_name %in% taster_sample,
           country %in% country_sample
         )

# let's check if done correctly:
wine_sample %>% 
  distinct(country) %>% 
  nrow() #5, correct!

wine_sample %>% 
  distinct(taster_name) %>% 
  nrow() #3, suspicious, but maybe only 3 tasters tasted wine from our subsample 
# of 10 countries

wine_sample

```

## Question 9
Using your wine_sample, create a new column difference_from_avg which contains 
the difference between each wine’s rating (points) and the average rating for 
that taster, then arrange the differences to see the highest positive difference first. 

```{r}
wine_sample %>% 
  group_by(taster_name) %>% 
  mutate(average_points = mean(points)) %>% 
  mutate(difference_from_avg = points - average_points) %>% 
  arrange(desc(difference_from_avg)
          )

```

