---
title: "R Notebook"
output: html_notebook
---

# Homework week 11 - day2

```{r}
library(rpart)
library(rpart.plot)
library(tidyverse)
library(GGally)

library(tidyverse)
titanic_set <- read_csv('data/titanic_decision_tree_data.csv')

shuffle_index <- sample(1:nrow(titanic_set))

# shuffle the data so class order isn't in order - need this for training/testing split later on 
titanic_set <- titanic_set[shuffle_index, ]
```

## Question 1


```{r}
#cleaning
titanic_tidy <- titanic_set %>% 
  filter(!is.na(survived)) %>% 
  mutate(sex = as_factor(sex),
         survived_flag = as_factor(if_else(survived == 1, TRUE, FALSE)),
         pclass = as_factor(pclass),
         embarked = as_factor(embarked)) %>% 
  mutate(age_status = as_factor(if_else(age > 16, "adult", "child"))) %>% 
  filter(!is.na(age_status)) %>% 
  select(-...1, -passenger_id, -name, -ticket, -fare, -cabin, -age, -survived)

titanic_tidy
```

## Question 2

Have a look at your data and create some plots to ensure you know what you’re 
working with before you begin. Write a summary of what you have found in your 
plots. Which variables do you think might be useful to predict whether or not 
people are going to die? Knowing this before you start is the best way to have 
a sanity check that your model is doing a good job.

```{r message = FALSE}
titanic_tidy %>% 
ggpairs()

```

Based on the GGpairs plot, looking at survived column it looks like:
sex -> gives a good indication if someone will survive
age_status -> good indication
pclass -> gives a reasonable indication 

Not great indicators are:  embarked, parch, and sib_sp

## Question 3

80% training, 20% saved for testing

```{r}
# how many rows in total
n_data <- nrow(titanic_tidy)
#714

proportion <- 0.2

# create a test sample index
test_index <- sample(1:n_data, size = n_data*proportion)

# split the data
titanic_test <- slice(titanic_tidy, test_index)
titanic_train <- slice(titanic_tidy, -test_index)


#check if balanced:

titanic_test %>% 
  janitor::tabyl(survived_flag)

titanic_train %>% 
  janitor::tabyl(survived_flag)

# proportions very close, so all ok!
```

## Question 4

Create your decision tree to try and predict survival probability using an 
appropriate method, and create a decision tree plot.

```{r}
# 1. make a tree model
titanic_fit <- rpart(
  formula = survived_flag ~ ., # include all variables
  data = titanic_train,
  method = "class" # 'class' for a categorical predictor
)

# 2. plot tree model
rpart.plot(titanic_fit,
           yesno = 2, # this writes y/n  at all splits
           type = 2, # dictates where our conditions lie at each node
           fallen.leaves = TRUE, # True means leaves all aligned at bottom
           faclen = 6, # length of factor names
           digits = 2, # how many decimal places is prob reported to
           split.border.col = 1 # colour of border factor names
           )
```

## Question 5

Write down what this tells you, in detail. What variables are important? 
What does each node tell you? Who has the highest chance of surviving? 
Who has the lowest? Provide as much detail as you can.

-> The decimal number gives the probability someone survived, given that they reach that leave
  While the proportion in % gives the total proportion of 'observations' at that point: the conditional probibility
  
Each node tells us:
  - predicted survival TRUE/FALSE at the top
  - decimal number gives us the probability of survival (of people meeting the conditions to reach this node)
  - % number, gives us the percentage of observations that pass through this node

So in short, the root gives us: 100% of the data/observations/passengers reach this point.
0.4 (=40%) meet the condition we are looking for (=survived_flag)

Highest chance of surviving:
0.93, with conditions: female in another pclass than 3

Lowest chance of survival:
0.06, with coditions: male, from pclass 2 or 3, child, with sib_sp less than 3

## Question 6

Test and add your predictions to your data. Create a confusion matrix. Write down 
in detail what this tells you for this specific dataset.

```{r}
#Use train model to add predictions to test data

library(modelr)

# add the predictions
titanic_test_pred <- titanic_test %>%
  add_predictions(titanic_fit, type = "class")

# confusion matrix
library(yardstick)

conf_mat <- titanic_test_pred %>% 
  conf_mat(truth = survived_flag,
           estimate = pred)

conf_mat
```
Confusion matrix tells us:
1. that for observations we predicted 'survival': 41 did in reality survive, 8 did not. 
  -> We have 8 false positives
2. of our predicted 'not-survival': 75 did not survive, and 18 did survive in reality.
  -> we have 18 false negatives
  
We can calculate the Sensitivity (true positive rate)
```{r}
sensitivity <- titanic_test_pred %>% 
  sensitivity(truth = survived_flag, estimate = pred)

sensitivity
#0.9 good!
```

We can calculate the Specificity (true negative rate)

```{r}
specificity <- titanic_test_pred %>% 
  specificity(truth = survived_flag, estimate = pred)

specificity
#0.69 less good but ok..
```










