---
title: "R Notebook"
output: html_notebook
---

# Homework Week 05 - Day 1: Introductory ggplot2 Homework

```{r}
# data is in the codeclan data package

library(tidyverse)
library(CodeClanData)
```
```{r}
#following data required:
view(qb_revenue_breakdown)
view(qb_monthly_sales)
view(qb_competitors)
view(qb_device_data)
```

## Question 1 - 
Take the data in the dataset `qb_revenue_breakdown` and make a stacked bar chart 
showing the sources of revenue across the two years in the dataset.

```{r}
ggplot(qb_revenue_breakdown) +
  aes(x = Year, y = Revenue, fill = Product) +
  geom_col()
```

## Question 2 -
Make a line chart showing monthly sales of the "flagship product" - the 'KwikBit Unit 1' - 
over the last year. This data is inside `qb_monthly_sales`, but you will need to 
use `filter()` or `subset()` (base `R)`.

```{r}
kwikbit_sales <- qb_monthly_sales %>% 
  filter(Cashflow == "Kwikbit Unit 1 Sales")

ggplot(kwikbit_sales) +
  aes(x = Date, y = Amount) +
  geom_line() +
  ggtitle("Monthly sales Kwikbit unit 1")


```

## Question 3 - 
Make a line chart showing monthly revenue and costs over the last year. 
This data is also in `qb_monthly_sales`.

```{r}
qb_monthly_sales %>% 
  filter(Cashflow != "Kwikbit Unit 1 Sales") %>%  #first filter, than pipe into ggplot

ggplot() +
  aes(x = Date, y = Amount, colour = Cashflow) +
  geom_line()
```

## Question 4 - 
Show annual sales of personal fitness trackers over the last 5 years broken down 
by company as a ribbon plot (use `geom_area`). This data is in `qb_competitors`.

```{r}
ggplot(qb_competitors) +
  aes(x = Year, y = Revenue, fill = Company) +
  geom_area()

# I think the revenue can be easier to read after transformation:

qb_competitors %>% 
  mutate(Revenue_thousands = Revenue/1000) %>% 
  ggplot() +
  aes(x = Year, y = Revenue_thousands, fill = Company) +
  geom_area() +
  ylab("Revenue (x1000)")

```

## Question 5 -
Now show the sales from the four competitors as a line graph. Include an extra layer 
that shows the data points used to make the lines.

```{r}
qb_competitors %>% 
  mutate(Revenue_thousands = Revenue/1000) %>% 
  ggplot() +
  aes(x = Year, y = Revenue_thousands, colour = Company) +
  geom_line() +
  ylab("Revenue (x1000)") +
  geom_point() #adding poiints to the lines
```

## Question 6 -
Now the company wants to compare the number of steps that their device counts vs. their competitors.
Make a line graph of the number of steps throughout time, and use faceting to compare between 
companies and people. The data for this is in `qb_device_data`.

```{r}
# exploring data
  #names(qb_device_data)
  #glimpse(qb_device_data)

    #qb_device_data %>% 
    #   distinct(device)
qb_device_data %>% 
  mutate(total_hours = (hours+(mins/60))) %>% 
  ggplot() +
  aes(x = total_hours, y = counts, group = id, colour = device) +
  geom_line() +
  facet_wrap(~ device, nrow = 1)

# Based on the answers, I am missing 

qb_device_data %>% 
  mutate(total_hours = (hours+(mins/60))) %>%
  ggplot() +
  aes(x = total_hours, y = counts) +
  geom_line() +
  facet_grid(id ~ device)

```

## Part 2: Question 1
Take the plots that you produced in part one and now polish them by:

1. Adding appropriate labels
2. Changing the scales and coordinates when appropriate.
3. Applying a unified theme, which is described below:
  * Graphs should have white backgrounds, and use colour sparingly.
  * There should be faint grid lines.
  * Font sizes should be ~12pt although titles should be slightly larger and axis labels can be slightly smaller. 
  * All plots should use colours from the following company colour scheme.
  col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")

```{r}
# Stacked bar chart
ggplot(qb_revenue_breakdown) +
  aes(x = Year, y = Revenue, fill = Product) +
  geom_col() +
  ylab("Revenue (£)\n") +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("Sources of revenue across years\n") +
  theme_light() +
  scale_fill_manual(
    values = c(
      "Accessories" = "#E89FE9",
      "Smart Watches" = "#50434F",
      "Subscription Services" = "#B6A7B5",
      "Trackers" = "#F9A472"
    ))

```

```{r}
# line chart monthly sale

kwikbit_sales <- qb_monthly_sales %>% 
  filter(Cashflow == "Kwikbit Unit 1 Sales")

ggplot(kwikbit_sales) +
  aes(x = Date, y = Amount) +
  geom_line(colour = "#E89FE9") +
  ggtitle("Monthly sales Kwikbit unit 1\n") +
  ylab("Number of units sold\n") +
  xlab(NULL) +
  theme_light()


```

```{r}
# line chart, revenue and costs

qb_monthly_sales %>% 
  filter(Cashflow != "Kwikbit Unit 1 Sales") %>%  #first filter, than pipe into ggplot

ggplot() +
  aes(x = Date, y = Amount, colour = Cashflow) +
  geom_line() +
  ggtitle("Qikbit revenue and costs") +
  ylab("Amount (£)\n") +
  theme_light() +
  xlab(NULL) +
  scale_color_manual(
    values = c(
      "Cost" = "#B6A7B5",
      "Total Revenue" = "#F9A472"
    )
  )

```

```{r}
# annual sales of competitors
qb_competitors %>% 
  mutate(Revenue_thousands = Revenue/1000) %>% 
  ggplot() +
  aes(x = Year, y = Revenue_thousands, fill = Company) +
  geom_area() +
  ggtitle("Revenue of competitors\n") +
  ylab("Revenue (x1000)\n") +
  theme_light() +
  scale_fill_manual(
    values = c(
      "Competitor 1" = "#E89FE9",
      "Competitor 2" = "#50434F",
      "Competitor 3" = "#B6A7B5",
      "Competitor 4" = "#F9A472"
    ))

```
```{r}
# annual sales competitors as line graph
qb_competitors %>% 
  mutate(Revenue_thousands = Revenue/1000) %>% 
  ggplot() +
  aes(x = Year, y = Revenue_thousands, colour = Company) +
  geom_line() +
  ylab("Revenue (x1000)\n") +
  geom_point() +
  theme_light() +
  ggtitle("Revenue of competitors\n") +
  scale_colour_manual(
    values = c(
      "Competitor 1" = "#E89FE9",
      "Competitor 2" = "#50434F",
      "Competitor 3" = "#B6A7B5",
      "Competitor 4" = "#F9A472"
    ))
```
```{r}

qb_device_data %>% 
  mutate(total_hours = (hours+(mins/60))) %>% 
  ggplot() +
  aes(x = total_hours, y = counts, group = id, colour = device) +
  geom_line() +
  theme_light() +
  ylab("Steps counted\n") +
  xlab("Hours counted") +
  labs(colour = "Product") +
  ggtitle("Number of steps counted competitors vs. qikbit 1\n") +
  facet_wrap(~ device, nrow = 1) +
  scale_colour_manual(
    values = c(
      "qikbit1" = "#E89FE9",
      "competitor1" = "#50434F",
      "competitor2" = "#B6A7B5",
      "competitor3" = "#F9A472"
    ))

# Based on the answers, I am missing 
col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")

theme_qb <- 
  theme(
    text = element_text(size = 12),
    title = element_text(size = 14),
    axis.text = element_text(size = 10),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "grey90", linetype = "dashed")
  )

#trying to edit my code based on answers
qb_device_data %>% 
  mutate(total_hours = (hours+(mins/60))) %>% 
  ggplot() +
  aes(x = total_hours, y = counts, colour = device) + #don't forget the colour=device, otherwise it does not know what to colour!!
  geom_line() +
  facet_grid(id ~ device) +
  ggtitle("Number of steps counted competitors vs. qikbit 1\n") +
  ylab("Steps counted\n") +
  xlab("Hours counted") +
  scale_colour_manual(values = col_scheme) +
  theme_qb


#From the answers:

qb_device_data <- 
mutate(qb_device_data, decimal_hour = hours + (mins/60))

ggplot(qb_device_data) +
  aes(x = decimal_hour, y = counts, colour = device) +
  geom_line(show.legend = FALSE) + 
  scale_x_continuous("Time (hours)") +
  scale_y_continuous("Steps") +
  facet_grid(id ~ device) +
  scale_colour_manual(values = col_scheme) +
  theme_qb +
  ggtitle("Comparison between KwikBit Unit1 and Competitors for 5 individuals")
```

```{r}
#Further useful:
scale_x_date(date_breaks = "1 month", date_labels = "%d-%b-%y") # (to split dates by month in the labels)



```

