---
title: "R Notebook"
output: html_notebook
---

# Homework Week 6 - day 1

## 1.1 The titanic
Calculate the conditional probabilities prob(survived∣crew) and prob(survived∣passenger)
 to settle this dispute.
 
```{r}
library(tidyverse)
library(gtools)
source("prob.R")

tit <- read_csv("data/titanic.csv")
```
```{r}
head(tit)

# lets change the data so its easier to calculate
tit %>% 
  distinct(class)

crew_tit <- tit %>% 
  select(class, survived) %>% 
  mutate(crew = class %in% c("engineering crew", "victualling crew", "restaurant staff", "deck crew")) 
```
 We need to calculate:
 prob(survived ∣ crew) = prob(survived and crew) / prob(crew)
 
```{r}
#dplyr method:

#----prob(survived ∣ crew)------
#prob(survived and crew)
crew_tit %>% 
  filter(survived == "yes", crew == TRUE) %>% 
  summarise(p = n()/nrow(crew_tit))
# = 0.09560489	

#prob(crew)
crew_tit %>% 
  filter(crew == TRUE) %>% 
  summarise(p = n()/nrow(crew_tit))
# = 0.4032623	
  
prob_survived_crew <- 0.09560489 / 0.4032623
prob_survived_crew # prob(survived ∣ crew) = 0.237

#----prob(survived∣passenger)------
#prob(survived and passenger)
crew_tit %>% 
  filter(survived == "yes", crew == FALSE) %>% 
  summarise(p = n()/nrow(crew_tit))
# = 0.2265519	

#prob(passenger)
crew_tit %>% 
  filter(crew == FALSE) %>% 
  summarise(p = n()/nrow(crew_tit))
# = 0.5967377

prob_survived_passenger <- 0.2265519 / 0.5967377
prob_survived_passenger #prob(survived∣passenger) = 0.379

```
Answer:
prob(survived∣passenger) = 0.379
prob(survived ∣ crew) = 0.237

passengers were more likely to survive!

```{r}
# you could also have used the contingency table:
library(janitor)

crew_tit %>% 
  tabyl(survived, crew) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages(denominator = "all")
```

## 1.2 Probability Consolidation Questions
 
### Question 1
If I roll a six-sided dice once, what is the probability of getting a number less than 5?
```{r}

#4 possibilities of throwing less than 5
(1/6)^4 #0.0007716049

#Based on answers, stupid mistake:
4/6 # -> 0.6667

```
 
### Question 2
How many outcomes are there in the sample space of an experiment in which we toss three dice?
```{r}
rolldie(3, 6) %>% 
  nrow()

# 216 different outcomes
# or 6*6*6


```

### Question 3
In an experiment where we roll four dice:

how many outcomes are in the event where the total of the four dice equals 5, 6 or 7
)?
what is the probability of this event?
what assumptions do we apply in answering part ii?

```{r}
rolldie(4, 6) %>% 
  mutate(total = X1+X2+X3) %>% 
  filter(total == 5 | total == 6 | total == 7) %>% 
  nrow()
# total 186 outcomes where total is 5, 6 or 7 

# probability of this event is:
rolldie(4, 6) %>% 
  summarise(p = 186/n())
# 0.1435185

# We assume that we use 4 independent dice
```

Based on answers, was wrong here (forgot to add X4:
```{r}
rolldie(4, 6) %>% 
  mutate(total = X1+X2+X3+X4) %>% 
  filter(total == 5 | total == 6 | total == 7) %>% 
  nrow()
# 34 possibilities

# probability of this event is:
rolldie(4, 6) %>% 
  summarise(p = 34/n())
# 0.0262

# We assume that we use 4 independent dice
```


### Question 4
Five chairs are placed in a row in a room, along with five people of differing height. 
The people sit down randomly on the chairs. What is the probability that they 
sit in order of decreasing height, left to right?

```{r}
# In total there are 120 different possibilities
permutations(5, r = 5, repeats.allowed = FALSE) %>% 
  nrow()
# or: 5! (5*4*3*2*1)

# each time only a single person is the largest, and this person has to sit on one particular chair

# this is a permutation
# p = N! / (N - r)!

# So here:
# 5*4*3*2*1/ (5*4*3*2*1 - 5*4*3*2*1) = 5*4*3*2*1/5*4*3*2*1 = 1

5*4*3*2*1
# answer:
1/120
#0.008333333
```

### Question 5
I roll six dice:

what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the 
dice, left to right? what is the probability I see one each of the numbers 
6, 5, 4, 3, 2 and 1 on the dice regardless of order?

```{r}
rolldie(6, 6) %>% 
  nrow() # 46656 options in total

# permutations ()
permutations(6, r = 6, repeats.allowed = FALSE) %>% 
  nrow()
# 720
# 720/46656 = 0.0154321


# combinations ()
combinations(6, 6, repeats.allowed = FALSE) %>% 
  nrow()
# 1
# (1/6)^6 # or 1/46656 #= p = 2.143347e-05

#?? there is something not right:

# all dice on order (only one occurance) = a permutation
# p = N! / (N - r)!
# (6*5*4*3*2*1) / ((6*5*4*3*2*1)-((6*5*4*3*2*1))
((6*5*4*3*2*1) / ((6*5*4*3*2*1)-((6*5*4*3*2*1))) = 1
  # or:
  (1/6)^6 # or 1/46656 #= p = 2.143347e-05

# all dice on order (no particular occurance) = 
#total_n_combinations = N! / ((N - r)! * r!) = 
  
  
#############
#  ----------Second attempt ------------
 # Second attempt
#1. all dice on order (only one occurance) 
 (1/6)^6 # or 1/46656 #= p = 2.143347e-05
  
#2. all dice numbered but not in order:
  (6/6)*(5/6)*(4/6)*(3/6)*(2/6)*(1/6) # = 0.0154321

  
```

### Question 6
If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades? 
```{r}
# spades is black
# there are 26 cards black in the game
# of 26, 13 are spades

13/26 # -> 1/2 chance its spades

```

